Pipit on the Post: Proving Pre- and Post-conditions of Reactive Systems

### 2024-01-08 2024pipit/section/40-Transition.tex
ok

>>>>

\TODO{describe translation}
The substitution-based semantics is better for equational reasoning, and (subjectively perhaps) simpler.
The transition system is better for certain automated proofs: a surprising number of interesting properties can be proven with k-induction alone.

\subsection{Translation correctness proofs}

\TODO{describe the proofs of translation correctness}
The translation to a transition system above is an abstraction.
We prove it.
Interesting lemmas etc.

### 2024-01-08 2024pipit/section/50-Extraction.tex
ok

>>>>

\TODO{}
\begin{itemize}
  \item C codegen: note about duplicating work, sketch potential solution;
  \item C codegen: future work: sketch potential solution for separate compilation, compare to Accelerate FFI \cite{clifton2014embedding};
  \item staging: similarity to \cite{gallego2021w};
  \item this extraction is very similar to the transition system translation, except translating contracts uses the body and ignores the rely and guarantee;
  \item talk about proof of correctness
\end{itemize}

### 2024-01-08 2024pipit/section/70-Related.tex
ok

\section{Related work}
\label{s:related-work}
% TODO arrays

@@

\subsection{Counterexamples and invariant generation}
\TODO{integration with model-checker}

>>>>

\section{Related work}
\label{s:related-work}

% TODO arrays
\TODO{requires rewrite}

@@

\subsection{Model-checking, counterexamples and invariant generation}
\TODO{integration with model-checker}
Model-checkers do have distinct advantages over interactive theorem-provers: they are more usable, can find strengthening invariants automatically, and provide counterexamples.
However, they are limited to comparatively simple properties.
The time-triggered example from \autoref{s:motivation} would be difficult to encode in a model-checker with a limited specification language, as the trigger invariant uses an auxiliary recursive function to find the next active trigger.

In future work, we would like to explore integrating Pipit with an existing model-checker.
For simple first-order nonrecursive programs and specifications that fit within the bounds of another model-checker, we could generate an encoding to use for counterexample generation.
We would also like to investigate automatically lifting the strengthened invariants and suggesting the corresponding Pipit invariant.
Such a translation and lifting would not form part of the trusted computing base, as generated invariants would still be proved separately in \fstar{}.
In this way, we hope that it may be possible to obtain some of the usability benefits of model-checkers, with the added power of manual proofs when necessary, while retaining trust that the compiled code satisfies the original specification.
